---
title: "Project"
author: "Keyan Chen"
date: "`r Sys.Date()`"
output: 
  html_document:
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(dpi = 600,
                      echo = TRUE, 
                      fig.keep = TRUE,
                      cache = TRUE)
```

```{r}
sessionInfo()
```

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(lubridate)
library(factoextra)
library(gridExtra)
library(mltools)
library(tmaptools)

set.seed(729)
```

```{r message=FALSE, warning=FALSE}
my.data <- read_csv("Airplane_Crashes_and_Fatalities_Since_1908.csv") %>% 
  tibble()
```

# Data Wrangle

```{r}
head(my.data)
tail(my.data)
```

First, we need to deal with the missing values. 

```{r}
my.data$Time[is.na(my.data$Time)] <- '00:00'
my.data$Time <- str_replace_all(my.data$Time, 'c: ', '')
my.data$Time <- str_replace_all(my.data$Time, 'c:', '')
my.data$Time <- str_replace_all(my.data$Time, 'c', '')
my.data$Time <- str_replace_all(my.data$Time, '\'', ':')
my.data$Time <- str_replace_all(my.data$Time, '\\.', ':')
my.data$Time[!is.na(str_match(my.data$Time, '[0-9]{4}'))]
```

Modify the time `0943`. 

```{r}
my.data$Time <- str_replace_all(my.data$Time, '0943', '09:43')
```

```{r}
my.data$Time[!is.na(str_match(my.data$Time, '[0-9]{3}'))]
my.data$Time <- str_replace_all(my.data$Time, '114:20', '00:00')
```

Modify `114:20`. 

```{r}
my.data$Time <- paste(my.data$Date, my.data$Time)
my.data$Time <- mdy_hm(my.data$Time)
my.data <- my.data %>%
  select(-c(`Flight #`, Route, Registration, `cn/In`, Summary)) %>%
  na.omit()
```

# Visualizing

```{r time_flow}
my.data %>%
  group_by(year(Time)) %>%
  summarise(count = n()) %>%
  ggplot(aes(`year(Time)`, count)) +
  geom_line(color = "gray52") +
  geom_point(color = "sienna1") +
  theme_light() +
  xlab("Year") +
  ylab("Counts")
```

The number of air crashes was higher from 1940 to 2005. Next, let's see if the crash is related to a specific number of months, days of the week, or a specific time period. 

```{r}
p.week <- my.data %>%
  group_by(wday(Time, week_start = 1, label = TRUE)) %>%
  summarise(count = n()) %>%
  ggplot(aes(`wday(Time, week_start = 1, label = TRUE)`, count)) +
  geom_col(fill = "#90AFC5", width = 0.8) +
  theme_light() +
  xlab("Day of Week") +
  ylab("Counts")
```

```{r}
p.month <- my.data %>%
  group_by(month(Time, label = TRUE)) %>%
  summarise(count = n()) %>%
  ggplot(aes(`month(Time, label = TRUE)`, count)) +
  geom_col(fill = "#336B87", width = 0.8) +
  theme_light() +
  xlab("Month") +
  ylab("Counts")
```

```{r}
p.day <- my.data %>%
  filter(hour(Time) != 0 | minute(Time) != 0) %>%
  group_by(hour(Time)) %>%
  summarise(count = n()) %>%
  ggplot(aes(`hour(Time)`, count)) +
  geom_col(fill = "#2A3132", width = 0.8) +
  theme_light() +
  xlab("Hour") +
  ylab("Counts") +
  scale_x_continuous(breaks = seq(0, 24, 4))
```

A significant increase in the number of air crashes can be seen during certain time periods. But we don't know the exact reason for this increase. 

It is possible that some factors during these specific periods of time have led to a higher probability of air crashes. It could also be that the number of flights increased during the specific time period, which led to an increase in the number of crashes.


```{r message=FALSE, warning=FALSE}
# library(Rmisc)
```


```{r Time_vs_Counts}
# multiplot(p.week, p.month, p.day, layout = matrix(c(1, 2, 3, 3),
#                                                   nrow = 2, byrow = TRUE))
```

```{r}
my.data <- my.data %>%
  mutate(Survivors = Aboard - Fatalities,
         SurvivalRate = Survivors / Aboard) %>% 
  na.omit()
```

```{r}
summary(my.data)
```
```{r}
top_n(count(my.data, Location, sort = TRUE), 10)
```
```{r}
top_n(count(my.data, Operator, sort = TRUE), 10)
```
```{r}
top_n(count(my.data, Type, sort = TRUE), 10)
```
```{r time_vs_surrate}
temp.data <- my.data %>%
  group_by(year(Time)) %>%
  summarise(Aboard = sum(Aboard),
            Fatalities = sum(Fatalities)) %>%
  ungroup() %>%
  mutate(Sur = Aboard - Fatalities,
         SurRate = Sur / Aboard) %>%
  na.omit()

temp.data %>%
  ggplot(aes(`year(Time)`, SurRate)) +
  geom_col(fill = 'gray60') +
  theme_light() +
  xlab('Years') +
  ylab('Survival Rate') +
  geom_hline(yintercept = 0.166) +
  geom_text(
    aes(x = 1908, y = 0.166),
    label = 'Average',
    vjust = 1.5,
    hjust = 0.5,
    color = 'indianred1'
  )
```


```{r}
cluster.data <- my.data %>%
  select(c(6:7, 9)) %>%
  scale()
```


```{r}
k2 <- kmeans(cluster.data, centers = 2, nstart = 25)
k3 <- kmeans(cluster.data, centers = 3, nstart = 25)
k4 <- kmeans(cluster.data, centers = 4, nstart = 25)
k5 <- kmeans(cluster.data, centers = 5, nstart = 25)
```


```{r k-means results}
p2 <-
  fviz_cluster(k2, geom = "point", data = cluster.data) + ggtitle("K-means k=2") +
  theme_light()
p3 <-
  fviz_cluster(k3, geom = "point", data = cluster.data) + ggtitle("K-means k=3") +
  theme_light()
p4 <-
  fviz_cluster(k4, geom = "point", data = cluster.data) + ggtitle("K-means k=4") +
  theme_light()
p5 <-
  fviz_cluster(k5, geom = "point", data = cluster.data) + ggtitle("K-means k=5") +
  theme_light()

grid.arrange(p2, p3, p4, p5)
```
```{r elbow}
fviz_nbclust(cluster.data, kmeans, method = "wss") +
  geom_vline(xintercept = 3, linetype = 2) +
  theme_light() +
  theme(title = element_blank())
```


```{r}
clustered.data <- tibble(my.data, cluster = k3$cluster)
```


```{r}
(summary_cluster <- clustered.data %>%
  group_by(cluster) %>%
  summarise(
    Counts = n(),
    Mean_Aboard = mean(Aboard),
    Mean_Fatalities = mean(Fatalities),
    Mean_Sur_Rate = mean(SurvivalRate)
  ))
```

```{r}
or <- summary_cluster$Mean_Sur_Rate %>% 
  rank()
```


```{r}
clustered.data$cluster_id <-
  ifelse(
    clustered.data$cluster == or[3],
    "Large Passenger High Survival",
    ifelse(
      clustered.data$cluster == or[1],
      "Large Passenger High Fatality",
      "Small-Midsize Crashes"
    )
  )
```


```{r}
LgSurvive <-
  clustered.data[clustered.data$cluster_id == "Large Passenger High Survival", ]
LgFatal <-
  clustered.data[clustered.data$cluster_id == "Large Passenger High Fatality", ]

Top10TypeSurvive <-
  top_n(count(LgSurvive, Type, sort = TRUE), 10)
Top10TypeFatal <-
  top_n(count(LgFatal, Type, sort = TRUE), 10)

Top10OperSurvive <-
  top_n(count(LgSurvive, Operator, sort = TRUE), 10)
Top10OperFatal <- top_n(count(LgFatal, Operator, sort = TRUE), 10)
```


```{r Type10_1}
Top10TypeSurvive %>%
  ggplot(aes(x = reorder(Type, n), y = n)) +
  geom_col(fill = "mediumseagreen") +
  coord_flip() +
  # ggtitle("Top 10 Large Passenger High Survival Crashes by Model") +
  xlab("Plane Model") +
  ylab("Number of Crashes") +
  theme_light()
```


```{r Type10_2}
Top10TypeFatal %>%
  ggplot(aes(x = reorder(Type, n), y = n)) +
  geom_col(fill = "orangered3") +
  coord_flip() +
  # ggtitle("Top 10 Large Passenger High Fatality Crashes by Model") +
  xlab("Plane Model") +
  ylab("Number of Crashes") +
  theme_light()
```


```{r Op1}
ggplot(Top10OperSurvive, aes(x = reorder(Operator, n), y = n)) +
  geom_col(fill = "mediumseagreen") +
  coord_flip() +
  # ggtitle("Top Ten Large Passenger High Survival Crashes by Airline") +
  xlab("Operator") +
  ylab("Number of Crashes") +
  theme_light()
```


```{r Op2}
ggplot(Top10OperFatal, aes(x = reorder(Operator, n), y = n)) +
  geom_col(fill = "orangered3") +
  coord_flip() +
  # ggtitle("Top Ten Large Passenger High Fatality Crashes by Airline") +
  xlab("Operator") +
  ylab("Number of Crashes") +
  theme_light()
```

```{r}
Locations <-
  as.character(clustered.data$Location[clustered.data$cluster != 2])
Locations <- gsub("Near ", "", Locations)
Locations <- gsub("Off ", "", Locations)
Locations <- gsub("off ", "", Locations)
Locations <- gsub("Czechoslovakia", "Czechia", Locations)
Locations <- gsub("AFB", "", Locations)
Locations <- gsub("Llandow Airport, ", "", Locations)
Locations <- gsub("La Guardia ", "LaGuardia ", Locations)
Locations <- gsub("Staten Island / ", "", Locations)
Locations <- gsub("Morrocco", "Morocco", Locations)
Locations <- gsub("Guadaloupe", "Guadeloupe", Locations)
Locations <- gsub("USSR", "Russia", Locations)
Locations <- gsub("Rio de Janerio", "Rio de Janeiro", Locations)
Locations <- gsub("Papanga", "Pampanga", Locations)
Locations <- gsub("Island of ", "", Locations)
Locations <- gsub(", Yugoslavia", "", Locations)
Locations <- gsub("near Roussillon, ", "", Locations)
Locations <- gsub("Covington/Hebron, ", "", Locations)
Locations <- gsub(" (Namibia)", "", Locations)
Locations <- gsub("Morioko", "Morioka", Locations)
Locations <- gsub("Mt. Fuji, ", "", Locations)
Locations <- gsub("Yuhnov", "Yukhnov", Locations)
Locations <- gsub("Ukraine, Russia", "Ukraine", Locations)
Locations <- gsub("Staines, Surrey, ", "", Locations)
Locations <- gsub("Massachusett", "Massachusetts", Locations)
Locations <- gsub("Corunda", "Coruna", Locations)
Locations <- gsub("Thirty-five miles west of ", "", Locations)
Locations <- gsub("Mountains ", "", Locations)
Locations <- gsub(" near Kampung Ladang, ", "", Locations)
Locations <- gsub("Elburz Mtns., near ", "", Locations)
Locations <- gsub("New York, New York", "New York", Locations)
Locations <- gsub("300 nm NW of ", "", Locations)
Locations <- gsub("950 nm S of  ", "", Locations)
Locations <- gsub("Western PacifiOcean, ", "", Locations)
Locations <- gsub("PacifiOcean", "Pacific Ocean", Locations)
Locations <- gsub("AtlantiOcean", "Atlantic Ocean", Locations)
Locations <- gsub("Over the ", "", Locations)
Locations <- gsub("110 miles SW of ", "", Locations)
Locations <-
  gsub("DemocratiRepubliCogo",
       "Democratic Republic of Congo",
       Locations)
Locations <- gsub("Calilfornia", "California", Locations)
Locations <-
  gsub(", 570 miles northeast of Natal, Brazil", "", Locations)
Locations <-
  gsub(", 100 miles W of Galway Bay, Ireland", "", Locations)
Locations <- gsub(", 110 miles West of Ireland", "", Locations)
Locations <-
  gsub(", 116 miles WSW of Annette Island, Alaska", "", Locations)
Locations <- gsub("Mariana Islands", "", Locations)
Locations <- gsub("Atalayasa", "Atalaya", Locations)
Locations <- gsub("Wusterausen", "Wusterhausen", Locations)
Locations <- gsub("Vrastsa", "Vratsa", Locations)
Locations <- gsub("Dneprodzerzhinsk", "Kamianske", Locations)
Locations <- gsub("At ", "", Locations)
Locations <- gsub("near Ajaccio, ", "", Locations)
Locations <- gsub("En route Miami, FL - ", "", Locations)
Locations <- gsub("near Ueno Village, ", "", Locations)
Locations <- gsub("French Alps, ", "", Locations)
Locations <-
  gsub("Kimpo Air Base", "Gimpo International Airport", Locations)
Locations <- gsub("Carpich", "Carpish", Locations)
Locations <- gsub(" (Namibia)", "", Locations)
Locations <-
  gsub("Al Fujayrah", "Fujairah International Airport", Locations)
Locations <- gsub(", Kefallinia, Greece", "", Locations)
Locations <- gsub("Karatepe Mountains", "Karatepe", Locations)
Locations <- gsub("Jeddah, ", "", Locations)
Locations <- gsub("Laskarak, ", "", Locations)
Locations <- gsub(", Ustica, Italy, ", "", Locations)
Locations <- gsub("Southern Belarus", "Belarus", Locations)
Locations <- gsub("Syktyvar", "Syktyvkar", Locations)
Locations <- gsub("Ko Phuket", "Phuket", Locations)
Locations <- gsub("Krosnovodsk", "Krasnovodsk", Locations)
Locations <- gsub("Combi, ", "", Locations)
Locations <- gsub("Persian Gulf, near ", "", Locations)
Locations <- gsub("Armenia, ", "", Locations)
Locations <- gsub("Mt. Saint-Odile, near ", "", Locations)
Locations <- gsub("Mt. Lalaboy,  ", "", Locations)
Locations <- gsub("Khorog", "", Locations)
Locations <- gsub("Tidjika", "Tidjikja", Locations)
Locations <- gsub("Kahengula, ", "", Locations)
Locations <- gsub("Domincan", "Dominican", Locations)
Locations <- gsub("Charkhidadri", "Charkhi Dadri", Locations)
Locations <- gsub("Massachusettss", "Massachusetts", Locations)
Locations <- gsub("Amritsar, India / ", "", Locations)
Locations <- gsub("Republiof", "Republic of", Locations)
Locations <- gsub("Khorramabed", "Khorramabad", Locations)

# coords <- geocode_OSM(Locations, keep.unfound = TRUE)
# 
# saveRDS(coords, file = "coords.RData")

coords <- readRDS(file = "coords.RData")
```

```{r}
map.data <- clustered.data[clustered.data$cluster != 2, ]

map.data$Latitude <- coords$lat
map.data$Longitude <- coords$lon

map.data <- na.omit(map.data)
```

```{r map}
map_data("world") %>%
  ggplot() +
  coord_fixed() +
  geom_polygon(aes(x = long, y = lat, group = group),
               fill = "white",
               colour = "grey50") +
  geom_point(
    data = map.data,
    aes(
      x = Longitude,
      y = Latitude,
      color = cluster_id,
      size = 1 - SurvivalRate
    ),
    alpha = 0.2
  ) +
  theme_light() +
  # ggtitle("Large Passenger Plane Crash Locations 1933 - 2009") +
  scale_color_manual("Cluster", values = alpha(c("orangered3", "green"))) +
  theme(legend.position = "none") +
  scale_size(range = c(0, 2.5))
```

